A Mechanised Library for Generic (Co)Programming

### 2024-02-26 paper/paper.tex

In order to abstract recursion patters we need to be able to talk about functors
which have the suitable fixed-point properties. Instances of these functors are
the polynomial ones which are in one-to-one correspondence with the so called
\emph{containers}~\cite{AbbottAG05}.
In this section we introduce containers

\subsection{Functors and Containers}
\label{sec:containers}
In order to To construct fixpoints of functors, these need to build
\emph{strictly positive types}. A common approach to construct such strictly
positive endofunctors is to use \emph{container functors}~\cite{AbbottAG05}.
However, reasoning about container and container functor equality requires to
consider both functional extensionality and heterogeneous equality. This is
problematic since, as stated in Section~\ref{sec:intro}, our goals to remain
axiom-free.  Furthermore, standard representations of containers in Coq will
lead to OCaml code that contains \mintinline{OCaml}{Obj.magic}. We will explain
why, and our mechanisation of containers in the remainder of this section.

A container is defined by a type of \emph{shapes}, and a \emph{family of
position types} indexed by shapes.
\begin{minted}{Coq}
Context (Shape : Type) (Pos : Shape -> Type).
\end{minted}
An \emph{extension} of this container is a functor, whose action on objects is
given by:
\begin{minted}{Coq}
Record App (X : Type) := { shape : Shape; contents : Pos shape -> X }.
\end{minted}
In our mechanisation, we use \mintinline{Coq}{App} to represent container
extensions.  Generally, any functor that is isomorphic to a container extension
is called a \emph{container functor}. However, in our work we will call
\emph{container functor} to the functors defined by a specific container
extension.

>>>>

In order to abstract recursion patters we need to be able to talk about functors
which have the suitable fixed-point properties, i.e. those who have an initial
algebra. Instances of these functors are the polynomial ones.  A common approach
to construct such functors is to use \emph{container}~\cite{AbbottAG05}.  In
this section we introduce the basics of polynomial functors as containers
(Section~\ref{sec:containers}).  However, reasoning about containers equality
will require us to consider both functional extensionality and heterogeneous
equality. As stated in previous sections~\ref{sec:intro}, this is rather
problematic since a stric requirement for this work is that we should remain
axiom-free. For this purpose in Section~\ref{sec:types-morphisms} we introduce
a custom equivalence relation on types in order to avoid using the functional
extensionality axiom.

\subsection{Functors and Containers}
\label{sec:containers}
A container is defined by a type of \emph{shapes} and a \emph{family of position
types} indexed by shapes.
\begin{minted}{Coq}
Context (Shape : Type) (Pos : Shape -> Type).
\end{minted}
An \emph{extension} of this container is a functor defined as follows:
\begin{minted}{Coq}
Record App (X : Type) := { shape : Shape; contents : Pos shape -> X }.
\end{minted}
with the obvious action on morphisms. Generally, any functor that is isomorphic
to a container extension is called a \emph{container functor}. However, in our
work we will call \emph{container functor} to the functors defined by a specific
container extension.
